A quantitative framework for investigating the reliability of network construction

### 2018-01-16 manuscript/CI_interactions.tex
updated fig. caption, put xaxis ticks on top too
older changes
added Tomes' comments, abstract and first draft of conclusiong

@@

@@

    Representation of a species assemblage as a network offers a convenient summary of how the community is constructed as it simultaneously accounts for species composition and their interactions. A tabulation of the nodes (species) and of their relative abundances forms the basis for traditional metrics of community composition. On top of this information, interactions are represented by links between nodes so that networks provide additional, higher-order information on community structure. Empirical descriptions of ecological networks are, however, usually considered to be static representations of the communities and interactions they describe. This approach contrasts significantly with the widely recognized stochasticity of ecological communities \citep{Gotelli2000}. That is, whether the network is assembled based on aggregated data, a single intensive ``snapshot" sample, or expert knowledge, interactions are assumed to occur deterministically wherever and whenever the community is observed~\citep{Olesen2011a}.

    The assumption of static communities conflicts with known variation in both  the nodes and links included in a network. Community composition and species abundances vary from site to site~\citep{Baiser2012} and over time within a site~\citep{Olesen2011a}. Likewise, interactions vary over space~\citep{Kitching1987,Baiser2012}, time~\citep{Kitching1987,Olesen2011a}, and between individuals of a given species~\citep{Pires2011a,Fodrie2015,Novak2015}. Moreover, variability in community composition and interactions may or may not be closely related. The removal of a species from a site will obviously also remove its interactions, but the co-occurrence of potentially interacting species do not guarantee they will interact at a given place and time. Interactions can be lost, or not detected, if participating species remain present but do not co-occur temporally, are too rare to detect each other~\citep{Tylianakis2010}, because of environmental contingencies \citep{Poisot2015}, or through changes to individual preferences~\citep{Fodrie2015}.

    Beyond ``true" variation in network structure, several researchers have pointed to the importance of sampling intensity for the assessment of network structure (e.g.,~\citealp{Martinez1999,Bluthgen2006,Bluthgen2007}). Accumulation of interactions with increasing sampling effort suggests that it is even more challenging to document interactions than species richness \citep{Poisot2012}. As a result, it has been proposed that interactions should be described probabilistically and network metrics computed accordingly \citep{Poisot2016}. Despite this possibility, the acknowledgment that insufficient sampling effort limits our ability to describe variation in ecological networks has not yet led to a quantitative methodology to deal with the uncertainty generated by spatiotemporal variation in ecological interactions and by sampling. Moreover, even in extremely well-sampled networks uneven sampling across species (or pairs of species) can lead to the assumption that some species do not interact because they co-occur rarely or have not yet been observed co-occuring (see Box 1 for an example). It is therefore likely that nearly all network studies are missing some interactions and could benefit from an approach that acknowledges this uncertainty.

@@

  To define the problem associated with describing ecological interaction networks, we will start from the perspective of an empirical community ecologist faced with the task of describing an interaction network. This ecologist will be interested in generating a description of the species/nodes present, and the links between them~\citep{Roslin2016}.  Importantly, the information sought is conveyed by both the presence and \emph{absence} of links. The two types of information are not, however, equally certain. An observed link will always remain an observed link, but there are multiple reasons why a given link may not be observed. Thus, the detection of any interaction is a stochastic process. We consider that there are three nested levels of uncertainty contributing to this stochasticity:

@@

    First, and most fundamentally, we do not know whether or not a pair of species have the appropriate characteristics (or traits) to interact. We define the probability of an interaction $L$ given those characteristics $\mathbf{T}$ as $P(L | \mathbf{T})=\lambda$. Obviously, if $k$ (the number of observed interactions) is 0, it is possible that the two species cannot interact even if there were no time or environmental constraints preventing it from occurring. In such cases, $\lambda$ would take a value of 0. 
    % [[The previous sentence doesn't follow, k hasn't been defined and we don't know that it's zero. Also, are we assuming that these interactions have been observed over a large area/time like the Kopelke data? Otherwise couldn't something about that particular site prevent species interacting even if they otherwise could?]] 
    % [[With defining k I think this is fine... site uncertainty will come in later, and it's not necessary to assume anything about the sampling design to get from not observing two species to interact to suspecting that they don't (degree of confidence depending on the species involved). I've toned down "very likely" to "possible" though - does that help?]]
    But it is nonetheless possible that the interaction is a rare phenomenon that has never been documented before, with $\lambda>0$. This source of uncertainty is the one documented by trait-matching models~\citep{Bartomeus2016}. It arises because every model is imperfect and lacks information (i.e. about traits) defining the constraints on the interaction. In other words, with sufficient sampling and all information accessible, this interaction probability $\lambda$ should either tend to 0 or to 1 and the uncertainty vanishes. 
    % [[For species which actually co occur (?)]]
    % [[I think that gets covered under "sufficient sampling"]]

@@

    It could be that we know an interaction is feasible, i.e. $L=1$, but that it does not occur at a given location or moment in time because of local constraints such as inclement weather or lack of suitable habitat. We define the realisation of the interaction process with the variable $X$, given that the interaction is feasible, as a stochastic process with associated probability $P(X|L)=\chi$. This phenomenon of interaction contingencies is usually not considered in network studies, but there is a rich literature in community ecology about the contingencies of interactions. Phenological matching~\citep{MillerRushing2010,Gezon2016}, species preferences~\citep{Pires2011,Novak2015,Coux2016}, and fear effects of other species~\citep{} are just some of the factors contributing to variation in the frequency of interactions between a given pair of species.
    % We imagine for instance that an interaction between a gall and a parasitoid might not be recorded at a location because it was the wrong time of the year and the gall had not yet formed (phenological constraints), a late frost killed the larvae (an abiotic environmental constraint), another parasitoid species competitively excluded the parasitoid species of interest (a biotic environmental constraint) or simply the parasitoid is too rare to interact with all of its viable hosts (an abundance constraint).

@@

    Lastly, measurement errors are always a source of uncertainty in the observation of ecological processes. We define the detection of an interaction with the variable $D$, given an interaction is feasible and occurs in the local conditions, as a stochastic process with associated probability $P(D|X,L)=\delta$. Detection failure could happen for several reasons, for instance if the rearing of a parasitoid fails with inappropriate lab conditions, because of species mis-identification, or because the interaction is very rare (see~\citet{Wirta2014} for examples of some of these difficulties and partial solutions to them). Some sources of detection error could be minimised with appropriate sampling effort ($\delta$ will converge to one with increasing number of samples), but other sources are often difficult to reduce (e.g. the occurrence of cryptic species might require molecular analysis for appropriate taxonomic identification as in~\citet{Wirta2014,Frost2016}).

    % \subsection*{What is an observation of "no interaction"?}

    The combination of these three sources of uncertainty together leads to many potential explanations for the observation of an absence of interaction ($D$, $X$, and/or $L = 0$), but only the situation where $L = 0$ (a true absence) is relevant for the investigation of networks. While the observation of an interaction is straightforward to interpret, absence of interactions must be decomposed into different quantities. It is particularly important to rule out the situations where $D=0 \cup X = 1 \cup L=1$, i.e. the interaction occurred at the location but was not observed, and $D=1 \cup X = 0 \cup L =1$, the interaction is feasible and would have been detected but did not occur at the local site. Ecologists seek to measure the occurrence of true absences, which is the joint event $L=0 \cup X=1 \cup D=1$), but in reality they measure the marginal probability $P(L) = k/n$ where $k$ is again the number of observed interactions and $n$ the number of observed co-occurrences.

@@

  For the ecologist collecting or interpreting network data, the considerations above raise a major challenge: how may we infer whether unobserved interactions went undetected due to sampling, or whether they truly do not occur. How then may we refine our sampling approaches to reduce uncertainties, and to gain insights into the impact of multiple processes on field observations? Importantly, some sources of uncertainty can be minimised with appropriate sampling design and efforts while other sources are difficult or impossible to reduce since they are generated by chance variation created by the very process in which we are interested.

  The basic information that we are collecting consists of matrices of interaction partners. Our count of interactions between two species reflects uncertainty with respect to whether two species are capable of interacting when co-occurring at a given site ($\lambda$), chance variation in whether the interaction is realised ($\chi$),  and uncertainty with respect to whether it is detected when realised ($\delta$). In this case, what can we do to pin down any of these different types of variation?

  The obvious rule of thumb is to ``sample more" (see Fig.~\ref{upper_limits} for a demonstration of the power of increasing sample size), but we note that there are limits to the utility of this motto. Sampling more will clearly reduce uncertainty regarding the upper bound of the probability of interaction, and it will also increase the probability of detecting unlikely interactions (i.e. interactions occurring with a low probability). Yet, since the probability of observing the co-occurrence of two species will always be higher than the probability of observing their interaction (since the probability of interaction is conditional on both interaction partners being present; see Fig.~\ref{histograms}E-F), we will accumulate observations of co-occurrences faster than we will accumulate observations of interactions. Thus, the more we sample, the more zeros will appear in our interaction matrix, which may seem like a counter-intuitive outcome.

  As a solution towards deciding whether unobserved interactions were undetected due to sampling, or whether they truly do not occur,~\citet{Weinstein2017} used repeated sampling rounds to estimate the daily probability of detecting a hummingbird interaction, and to thereby model the observation and process mechanisms. While conceptually attractive, this approach is unsuitable for interactions occurring over longer time scales (e.g., associations between hosts and parasitoids with a single generation per year), or very rare interactions which might not occur on any of the sampling days or might involve individuals of a species that are not under observation. What is worse, the problem persists that if a given interaction is not observed on a given day, this could be either because it was impossible that day despite being otherwise feasible [$P(X|D=1,L=1)=0$], because interaction between the two did occur but could not be observed [$P(D|X=1,L=1)=0$], or from any combination of the two. From a conceptual perspective, this approach will thus fail to satisfactorily distinguish between sources of uncertainty. Most importantly, if two species are never observed co-occurring during several days of sampling then we have learned nothing about their probability of interacting if they should ever co-occur. In other words, there is no information about interactions without co-occurrence.

  An added complication is that not all sources of uncertainty are proportional to sample size. To record an interaction between A and B, we need to identify both partners correctly (a non-trivial problem in many food webs; e.g.~\citealp{Kaartinen2011,Roslin2016}) and be able to resolve all interactions with a similar likelihood. For both molecular and rearing techniques, certain types of interactions may go unnoticed due to technical challenges~\citep{Wirta2014}. This can bias the set of recorded interactions. The bottom line is that separating different sources of uncertainty is difficult indeed. As an alternative to abandoning empirical networks or continuing to ignore the uncertainty inherent in undetected observations, we propose that some insight regarding the detectability of interactions between species not found co-occurring may be gained from data on other species between which interactions are known to be feasible.

@@

  To progressively dissect the different contributions to uncertainty, we will start by considering how we could naively quantify interaction probability and its associated uncertainty \emph{for an interaction that has not yet been observed}. We consider that a pair of species have been observed $n$ times co-occurring at one location, and the number of times they have been observed interacting is $k = 0$. We now aim to evaluate the uncertainty of this interaction. The occurrence of an interaction is a Bernoulli trial. Consequently, the number of successes $k$ over $n$ trials will follow a binomial distribution:

@@

  The variance of a Bernoulli experiment is $n\lambda$(1-$\lambda$). It is important to remember that this variance describes the variability of the number of successes $k$ for $n$ trials and is not the variance associated with the estimation of $\lambda$. Given this variance, it is possible to compute the confidence interval for the MLE of $\lambda$ using any of several methods, including the \emph{Wilson score interval}, the \emph{Clopper-Pearson interval}, and the \emph{Agresti-Coull Interval} (for details, see [\citealp{}]). Finding this estimate is therefore quite straightforward, but it nonetheless has two drawbacks. First, $\lambda$ is not a single point estimate but rather a random variable with a given uncertainty, the distribution of which is unknown. This means that even if $k = 0$ in a given sample, this does not necessarily imply that the two species will never interact. What it implies is that no interaction is the most likely situation but there is nonetheless a substantial chance that the two species \emph{could} interact. In contrast with the situation where $k>0$, when we are sure that an interaction is feasible, we are not sure that absence of interactions is caused by $L=0$. Second, and most importantly, the number of trials $n$ might be very low in many instances (and some pairs of species might not have been documented as co-occuring even once). Consequently, the uncertainty in the estimation of $\lambda$ might be considerable. In Fig.~\ref{upper_limits} and Box~\ref{box2}, we derive the Clopper-Pearson interval to explore how the estimate of $\lambda$ varies with sample size. At a small sample size, the 95\% confidence interval spans all values of $\lambda$. To establish that species are not interacting with any acceptable certainty requires tens of observations of the two species co-occurring but not interacting. As most data sets will not have such extensive sampling across all observations, we can use a Bayesian approach to supplement what data we do have with other sources of information.

@@

      We here adopt a Bayesian approach to estimate the distribution of the parameter $\lambda$. According to Bayes principle, the posterior distribution of $\lambda$ is:

@@

     In some cases it can be complicated to compute the normaliser, but fortunately in our case an analytical solution exists. The beta-binomial distribution is a conjugate distribution of the binomial distribution. This allows us to analytically compute the posterior distribution of a binomial model with a beta prior distribution. We can re-write the posterior distribution of $\lambda$ as:

@@

          In the literature of ecological networks there is a solid collection of data for which we know network connectance and from which we can define the connectance distribution. Connectance is measured as $C = L/S^2$, where $L$ is the number of interactions and $S$ is the number of species. It measures the filling of an interaction matrix and thereby expresses the average probability that any two species interact with each other. If we know only the mean $\overline{C}$ and the variance $\sigma_C^2$ of the distribution of $C$,
          % [[from the literature or from our network? Has to be from the literature - if it was our network we'd know way more than just C]] 
          then the beta parameters could be computed as follows using the method of moments:

@@

          The degree of a node in a network is defined as the number of interactions it has to other nodes. The degree distribution of a network is then the probability distribution of these degrees over the whole network and the standardised degree could therefore be interpreted as an interaction probability. It is consequently possible to use the degree distribution to inform the prior distribution. The degree distribution could come from several networks, from a similar network (e.g. a known network at slightly different location) or from the network of interest if interaction probabilities for some species are already documented. The latter approach allows us to apply information from known, abundant species to the rarest species for which interactions are less frequently documented.

          If our focal network describes a similar system to that in a known network, we can use the distribution of interaction probabilities in that network to inform our prior. The probability of any interaction $L_ij$ depends on the degrees of species $i$ and $j$. Using normalised degrees $\Delta_i$ and $\Delta_j$ (i.e., degrees divided by the number of species in the network), we can obtain the probability of interaction $L_ij$=$\Delta_i\times\Delta_j$. Similar to the procedure for degree distribution, the distribution of these interaction probabilities derived from a similar network or a suite of networks can be used to establish a prior distribution before any data from the focal network are collected. For distributions of either degrees or interaction probabilities, the procedure for the estimation of the hyper parameters follows the same approach as described above for connectance except that each measurement is at the individual interaction level instead of the network level.

@@

          The last approach we envision to document the prior distribution of $\lambda$ is to use the outcome of a trait-matching model, provided such a model has been documented using external data and relevant traits are available. 
          In such as case, the prior distribution would follow the function $P(\lambda|\mathbf{T})=f(\mathbf{T})$ based on a set of traits for both species $\mathbf{T}$. There are several techniques available to perform this inference of interaction probability, some of which are Bayesian, and we refer to ~\citep{Bartomeus2016} and \citep{Weinstein2017} for recent reviews about this topic. Note that in this case the prior might not be beta distributed and numerical methods might be required to compute the posterior distribution.

@@

  The Bayesian framework can be illustrated with a simple quantitative example. Suppose we have $n = 10$ observations of co-occurrence between species $i$ and species $j$ in a given time interval and area, and $k = 3$ observations of interactions. The maximum likelihood estimate of the interaction probability (Box 1) is simply $\lambda_{MLE} = 3/10 = 0.3$.

@@

    Now, consider the case where the two species have never been observed interacting across $n$ trials, i-e. $k=0$. The question is then ``what is the probability these two species do not interact''? Since it is not possible to prove that the two species could never interact (strickly speaking, in this approach $\lambda=0$ is impossible), given the framework above we must fix a threshold below which we consider that there is no interaction ($\lambda \sim 0$). We call this threshold probability $\lambda*$. We then use the cumulative distribution function to estimate $P(\lambda<\lambda*|L=0,n)$ for different $n$. The function ``plot\_precision" in \emph{Appendix S2} calculates distribution function for $\lambda*$ with an increasing number of trials. This yields a surprising result: it requires \textgreater24 observations of no interactions to be 95\% sure that the interaction probability is smaller than $\lambda*$=0.1 (recall Fig.~\ref{upper_limits}, Box~\ref{box2}). Note the special case where there is no observation of the two species co-occurring, $n = 0$.  In this situation, the posterior distribution converges to the prior distribution since the data include no information on the probability with which species might interact should they co-occur).

@@

  [Here we need to be very hands-on and remember the invitation letter form Ecol Letters:
  As a contribution primarily to methodology, the manuscript will not be considered for ``Ideas \& Perspectives". (Please recall my message concerning your past presubmission inquiry.) However, Ecology Letters is entertaining the idea of a new section for ``Novel Methods". As we have not advertised this, I would ask you to keep that information in confidence. If acceptable to an Associate Editor and reviewers, your paper could be the first, or one of the first, contributions to this new section. Accordingly, the work would be subject to considerable scrutiny and I would urge you to make the paper as strong as you can. This would include demonstrating the need for this new method (i.e. outlining how this is different than other approaches currently in use and how they fail at some important task), documenting the robustness and accuracy of the approach, performing case study, and making robust high-quality code freely available.]

>>>>

% Main text can be up to 7500 words, 10 fig., tables, boxes

@@

  Current descriptions of ecological networks will typically assume that specific interactions occur wherever and whenever the community is observed. This runs in stark contrast with the widely recognized stochastic nature of ecological communities. When accurately resolved, community composition, species abundances and link structure have typically been found to vary in space and time. On top of this, finite sampling will generate variation in what links are observed and what not. Despite these challenges, we still lack the quantitative methodology to deal with the uncertainty generated by “true” spatiotemporal variation in ecological interactions and by sampling. In this study, we develop analytical tools to capture the uncertainty in the estimation of pairwise interactions. To define the problem, we first discuss the pervasive implications of uncertainty in the estimation of pairwise interactions on the properties of ecological networks. We then define a framework which identifies and dissects the different contributions to uncertainty of interaction. We illustrate this framework with a simple quantitative example using the most extensively sampled network to date. Next, we use the framework to provide analytical criteria for improving the sampling of rare interactions. Finally, we offer tangible recommendations for improved descriptors of ecological interactions. Through these efforts, we demonstrate both the utility of our approach and the importance of acknowledging the uncertainty inherent in network studies.

  % Can be up to 200 words

@@

    Representation of a species assemblage as a network offers a convenient summary of how the community is constructed as it simultaneously accounts for species composition and their interactions. A tabulation of the nodes (species) and of their relative abundances forms the basis for traditional metrics of community composition. On top of this information, interactions are represented by links between nodes so that networks provide additional, higher-order information on community structure. Empirical descriptions of ecological networks are, however, usually considered to be static representations of the communities and interactions they describe. That is, whether the network is assembled based on aggregated data, a single intensive ``snapshot" sample, or expert knowledge, interactions are assumed to occur deterministically wherever and whenever the community is observed~\citep{Olesen2011a}.

    The assumption of static communities contrasts significantly with the widely recognized stochasticity of ecological communities \citep{Gotelli2000}. Community composition and species abundances vary from site to site~\citep{Baiser2012} and over time within a site~\citep{Olesen2011a}. Likewise, interactions vary over space~\citep{Kitching1987,Baiser2012}, time~\citep{Kitching1987,Olesen2011a}, and between individuals of a given species~\citep{Pires2011a,Fodrie2015,Novak2015}. Moreover, variability in community composition and interactions may or may not be closely related. The removal of a species from a site will obviously also remove its interactions, but conversely, the co-occurrence of potentially interacting species does not in itself guarantee that they will interact at a given place and time. Interactions can be lost if the interaction partners remain present but are separated in time or are too rare to detect each other~\citep{Tylianakis2010}. Interactions can also fail to occur because of environmental contingencies \citep{Poisot2015}, or through changes to individual preferences~\citep{Fodrie2015}.

    Beyond ``true" variation in network structure, several researchers have pointed to the importance of sampling intensity for the assessment of network structure (e.g.,~\citealp{Martinez1999,Bluthgen2006,Bluthgen2007}). An assessment of the accumulation of interactions with increasing sampling effort suggests that it is even more challenging to document interactions than species richness \citep{Poisot2012}. As a result, it has been proposed that interactions should be described probabilistically and network metrics computed accordingly \citep{Poisot2016}. Yet, to dat, we lack the quantitative methodology to deal with the uncertainty generated by spatiotemporal variation in ecological interactions and by sampling. Moreover, even in extremely well-sampled networks uneven sampling across species (or pairs of species) can lead to the erroneous inference that some species do not interact because they co-occur rarely or have not yet been observed together - even if they do interact when they do co-occur (see Box 1 for an example). Nearly all network studies will thus neglect some interactions, necessitating an approach that acknowledges this uncertainty.

@@

  To define the problem associated with quantifying ecological interaction networks, we will start from the perspective of an empirical community ecologist faced with the task of describing a previously unknown interaction network. This ecologist will be interested in generating a description of the species/nodes present, and the links between them~\citep{Roslin2016}.  Importantly, the information sought is conveyed by both the presence and \emph{absence} of links. The two types of information are not, however, equally certain. An observed link will always remain an observed link, but there are multiple reasons why a given link may not be observed. Thus, the detection of any interaction is a stochastic process. We define three nested levels of uncertainty contributing to this stochasticity:

@@

    First, and most fundamentally, we do not know whether or not a pair of species have the appropriate characteristics (or traits) to interact. We define the probability of an interaction $L$ given those characteristics $\mathbf{T}$ as $P(L | \mathbf{T})=\lambda$. Obviously, if $k$ (the number of observed interactions) is 0, it is possible that the two species could not interact even if there were no external constraints (e.g., temporal or environmental separation) preventing the interaction from occurring. As a simple example, a prey species may be too large to be consumed by a particular predator. In such cases, $\lambda$ would take a value of 0. 
    % [[The previous sentence doesn't follow, k hasn't been defined and we don't know that it's zero. Also, are we assuming that these interactions have been observed over a large area/time like the Kopelke data? Otherwise couldn't something about that particular site prevent species interacting even if they otherwise could?]] 
    % [[With defining k I think this is fine... site uncertainty will come in later, and it's not necessary to assume anything about the sampling design to get from not observing two species to interact to suspecting that they don't (degree of confidence depending on the species involved). I've toned down "very likely" to "possible" though - does that help?]]
    Nevertheless, it is also possible that the interaction is a rare phenomenon with $\lambda>0$ that has not yet been documented. This source of uncertainty is the one documented by trait-matching models~\citep{Bartomeus2016}. It arises because every model is imperfect and lacks information (i.e. about traits) defining the constraints on the interaction. In other words, with sufficient sampling and all information accessible, this interaction probability $\lambda$ should either tend to 0 or to 1 and the uncertainty vanishes. 
    % [[For species which actually co occur (?)]]
    % [[I think that gets covered under "sufficient sampling"]]

@@

    Even when an interaction is feasible, i.e. $L=1$, it may not occur at a given location or moment in time because of local constraints such as inclement weather or the lack of suitable habitat. We define the realisation of the interaction process with the variable $X$, given that the interaction is feasible, as a stochastic process with associated probability $P(X|L)=\chi$. This phenomenon of interaction contingencies is usually not considered in network studies, but there is a rich literature in community ecology about the contingencies of interactions. Phenological matching~\citep{MillerRushing2010,Gezon2016}, species preferences~\citep{Pires2011,Novak2015,Coux2016}, and fear effects of other species~\citep{} are just some of the factors contributing to variation in the frequency of interactions between a given pair of species.
    % We imagine for instance that an interaction between a gall and a parasitoid might not be recorded at a location because it was the wrong time of the year and the gall had not yet formed (phenological constraints), a late frost killed the larvae (an abiotic environmental constraint), another parasitoid species competitively excluded the parasitoid species of interest (a biotic environmental constraint) or simply the parasitoid is too rare to interact with all of its viable hosts (an abundance constraint).

@@

    Lastly, measurement errors are a pervasive source of uncertainty in the observation of ecological processes. Given that an interaction with the variable $D$, given an interaction is feasible and occurs under the local conditions, we may define the detection of an interaction, $D$, as a stochastic process with the associated probability $P(D|X,L)=\delta$. Detection failure could happen for several reasons, for instance if the rearing of a parasitoid fails with inappropriate lab conditions, because of species mis-identification, or because the interaction is very rare (see~\citet{Wirta2014} for examples of some of these difficulties and partial solutions to them). Some sources of detection error could be minimised with appropriate sampling effort ($\delta$ will converge to one with increasing number of samples), but other sources are often difficult to reduce (e.g. the occurrence of cryptic species might require molecular analysis for appropriate taxonomic identification as in~\citet{Wirta2014,Frost2016}).

@@

    Together, the combination of these three sources of uncertainty --interaction uncertainty, process uncertainty, and detection uncertainty-- results in a range of potential explanations for the observation of an absence of interaction ($D$, $X$, and/or $L = 0$). The ecologist wanting to describe the network, however, is specifically interested in the situation where $L = 0$ (i.e., in true absences). Thus, while she or he will have no difficulty interpreting the observation of an interaction, the observation of an absence of an interaction offers more of a challence since it must be decomposed into different quantities. It is particularly important to rule out the situations where $D=0 \cup X = 1 \cup L=1$, i.e. where the interaction occurred at the location but was not observed, and $D=1 \cup X = 0 \cup L =1$, when the interaction is feasible and would have been detected but did not occur at the local site. The occurrence of a true absence, our quantity of interest, corresponds to the joint event $L=0 \cup X=1 \cup D=1$), but in reality an empirical ecologist will measure the marginal probability $P(L) = k/n$ where $k$ is again the number of observed interactions and $n$ the number of observed co-occurrences.

    The considerations above raise a major challenge: when faced with empirical data, how may we infer whether unobserved interactions went undetected due to sampling, or whether they truly do not occur. How then may we refine our sampling approaches to reduce uncertainties, and to gain insights into the impact of multiple processes on field observations? Importantly, some sources of uncertainty can be minimised with appropriate sampling design and efforts while other sources are difficult or impossible to reduce since they are generated by chance variation created by the very process in which we are interested.

  % [[Tomas suggests removing the following because it slows down procedings:]]
  % The basic information that we are collecting consists of matrices of interaction partners. Our count of interactions between two species reflects uncertainty with respect to whether two species are capable of interacting when co-occurring at a given site ($\lambda$), chance variation in whether the interaction is realised ($\chi$),  and uncertainty with respect to whether it is detected when realised ($\delta$). In this case, what can we do to pin down any of these different types of variation?

  % The obvious rule of thumb is to ``sample more" (see Fig.~\ref{upper_limits} for a demonstration of the power of increasing sample size), but we note that there are limits to the utility of this motto. Sampling more will clearly reduce uncertainty regarding the upper bound of the probability of interaction, and it will also increase the probability of detecting unlikely interactions (i.e. interactions occurring with a low probability). Yet, since the probability of observing the co-occurrence of two species will always be higher than the probability of observing their interaction (since the probability of interaction is conditional on both interaction partners being present; see Fig.~\ref{histograms}E-F), we will accumulate observations of co-occurrences faster than we will accumulate observations of interactions. Thus, the more we sample, the more zeros will appear in our interaction matrix, which may seem like a counter-intuitive outcome.

  % As a solution towards deciding whether unobserved interactions were undetected due to sampling, or whether they truly do not occur,~\citet{Weinstein2017} used repeated sampling rounds to estimate the daily probability of detecting a hummingbird interaction, and to thereby model the observation and process mechanisms. While conceptually attractive, this approach is unsuitable for interactions occurring over longer time scales (e.g., associations between hosts and parasitoids with a single generation per year), or very rare interactions which might not occur on any of the sampling days or might involve individuals of a species that are not under observation. What is worse, the problem persists that if a given interaction is not observed on a given day, this could be either because it was impossible that day despite being otherwise feasible [$P(X|D=1,L=1)=0$], because interaction between the two did occur but could not be observed [$P(D|X=1,L=1)=0$], or from any combination of the two. From a conceptual perspective, this approach will thus fail to satisfactorily distinguish between sources of uncertainty. Most importantly, if two species are never observed co-occurring during several days of sampling then we have learned nothing about their probability of interacting if they should ever co-occur. In other words, there is no information about interactions without co-occurrence.

  % An added complication is that not all sources of uncertainty are proportional to sample size. To record an interaction between A and B, we need to identify both partners correctly (a non-trivial problem in many food webs; e.g.~\citealp{Kaartinen2011,Roslin2016}) and be able to resolve all interactions with a similar likelihood. For both molecular and rearing techniques, certain types of interactions may go unnoticed due to technical challenges~\citep{Wirta2014}. This can bias the set of recorded interactions. The bottom line is that separating different sources of uncertainty is difficult indeed. As an alternative to abandoning empirical networks or continuing to ignore the uncertainty inherent in undetected observations, we propose that some insight regarding the detectability of interactions between species not found co-occurring may be gained from data on other species between which interactions are known to be feasible.

@@

  To progressively dissect the different contributions to uncertainty, we will start by considering how we could naively quantify interaction probability and its associated uncertainty \emph{for an interaction that has not yet been observed}. We consider the case where a pair of species have been observed co-occurring $n$ times, of which they have been observed to interact in $k = 0$ cases. We now aim to evaluate the uncertainty of this interaction. The occurrence of an interaction is a Bernoulli trial. Consequently, the number of successes $k$ over $n$ trials will follow a binomial distribution:

@@

  The variance of a Bernoulli experiment is $n\lambda$(1-$\lambda$). It is important to remember that this variance describes the variability of the number of successes $k$ for $n$ trials and is not the variance associated with the estimation of $\lambda$. Given this variance, it is possible to compute the confidence interval for the MLE of $\lambda$ using any of several methods, including the \emph{Wilson score interval}, the \emph{Clopper-Pearson interval}, and the \emph{Agresti-Coull Interval} (for details, see [\citealp{}]). Finding this estimate is therefore quite straightforward, but it nonetheless has two drawbacks. First, $\lambda$ is not a single point estimate but rather a random variable with a given uncertainty, the distribution of which is unknown. This means that even if $k = 0$ in a given sample, this does not necessarily imply that the two species will never interact. What it implies is that 'no interaction' is the most likely ø¨†come when the species do co-occur, but there is nonetheless a substantial chance that the two species \emph{could} interact. In contrast, in the situation where $k>0$, i.e., when we are sure that an interaction is feasible, we still cannot be sure that absence of interactions is caused by $L=0$. Second, and most importantly, the number of trials $n$ might be very low in many instances (and some pairs of species may never have been documented as co-occuring). Consequently, the uncertainty in the estimation of $\lambda$ might be considerable. In Fig.~\ref{upper_limits} and Box~\ref{box2}, we derive the Clopper-Pearson interval to explore how the estimate of $\lambda$ varies with sample size. At a small sample size, the 95\% confidence interval spans all values of $\lambda$. To establish that species are not interacting with any acceptable certainty requires tens of observations of the two species co-occurring but not interacting. As most data sets will lack such extensive sampling across all observations, we can use a Bayesian approach to supplement what data we do have with other sources of information.

@@

      Here we adopt a Bayesian approach to estimate the distribution of the parameter $\lambda$. According to Bayes principle, the posterior distribution of $\lambda$ is:

@@

     In some cases it can be complicated to compute the normaliser, but fortunately an analytical solution already exists in our case. The beta-binomial distribution is a conjugate distribution of the binomial distribution. This allows us to analytically compute the posterior distribution of a binomial model with a beta prior distribution. We can re-write the posterior distribution of $\lambda$ as:

@@

          In the literature of ecological networks there is a solid collection of networks for which connectance has been calculated and for which we can thus define the connectance distribution. Connectance is measured as $C = L/S^2$, where $L$ is the number of interactions and $S$ is the number of species. It measures the filling of an interaction matrix and thereby expresses the average probability that any two species interact with each other. If we know only the mean $\overline{C}$ and the variance $\sigma_C^2$ of the distribution of $C$,
          % [[from the literature or from our network? Has to be from the literature - if it was our network we'd know way more than just C]] 
          then the beta parameters could be computed as follows using the method of moments:

@@

          The degree of a node in a network is defined as its number of connections to other nodes. The degree distribution of a network is then the probability distribution of these degrees over the whole network and the standardised degree could therefore be interpreted as an interaction probability. It is consequently possible to use the degree distribution to inform the prior distribution. The degree distribution could come from several networks, from a similar network (e.g. a known network at slightly different location) or from the network of interest if interaction probabilities for some species are already documented. The latter approach allows us to apply information from known, abundant species to the rarest species for which interactions are less frequently documented.

          If our focal network describes a system similar to that in a known network, we can use the distribution of interaction probabilities in that network to inform our prior. The probability of any interaction $L_ij$ depends on the degrees of species $i$ and $j$. Using normalised degrees $\Delta_i$ and $\Delta_j$ (i.e., degrees divided by the number of species in the network), we can obtain the probability of interaction $L_ij$=$\Delta_i\times\Delta_j$. Similar to the procedure for degree distribution, the distribution of these interaction probabilities derived from a similar network or a suite of networks can be used to establish a prior distribution before any data from the focal network are collected. For distributions of either degrees or interaction probabilities, the procedure for the estimation of the hyper parameters follows the same approach as described above for connectance except that each measurement is at the individual interaction level instead of the network level.

@@

          As a fourth and final approach to obtain the prior distribution of $\lambda$, we envisage using the outcome of a trait-matching model, provided such a model has been documented using external data and relevant traits are available. 
          In such as case, the prior distribution would follow the function $P(\lambda|\mathbf{T})=f(\mathbf{T})$ based on a set of traits for both species $\mathbf{T}$. There are several techniques available to perform this inference of interaction probability, some of which are Bayesian, and we refer to ~\citep{Bartomeus2016} and \citep{Weinstein2017} for recent reviews about this topic. Note that in this case the prior might not be beta distributed and numerical methods might be required to compute the posterior distribution.

@@

  The Bayesian framework can be illustrated with a simple quantitative example. Suppose we have $n = 10$ observations of co-occurrence between species $i$ and species $j$ in a given time interval and area, and $k = 3$ observations of interactions. Then, the maximum likelihood estimate of the interaction probability (Box 1) is simply $\lambda_{MLE} = 3/10 = 0.3$.

@@

    Now, consider the case where the two species have never been observed interacting across $n$ trials, i-e. $k=0$. The question is then ``what is the probability that these two species do not interact''? Since it is not possible to prove that the two species could never interact (strictly speaking, in this approach $\lambda=0$ is impossible), we must fix a threshold below which we consider that there is no interaction ($\lambda \sim 0$). We call this threshold probability $\lambda*$. We then use the cumulative distribution function to estimate $P(\lambda<\lambda*|L=0,n)$ for different $n$. The function ``plot\_precision" in \emph{Appendix S2} calculates distribution function for $\lambda*$ with an increasing number of trials. This yields a surprising result: it requires \textgreater24 observations of no interactions to be 95\% sure that the interaction probability is smaller than $\lambda*$=0.1 (recall Fig.~\ref{upper_limits}, Box~\ref{box2}). Note the special case where there is no observation of the two species co-occurring, $n = 0$.  In this situation, the posterior distribution converges to the prior distribution since the data include no information on the probability with which species might interact should they co-occur.

@@

  Real interaction networks vary~\citep{}, and to capture this variation we must turn from static descriptions of network structure to probabilistic descriptions. In this study, we have developed the analytical tools to capture the uncertainty in the estimation of pairwise interactions, and to partition it into its individual components: interaction uncertainty, process uncertainty and detection uncertainty. The patters unveiled offer tangible recommendations for improved descriptors of ecological interactions. First, our analyses point to detection uncertainty as a major contributor to overall uncertainty of is establishing the absence of interaction. To counter these establish the true absence of interactions will then require comparatively large sample size – on the order of xxx. Second, xxxx. Third, xxxx. Through these efforts, we demonstrate both the utility of our approach and the importance of acknowledging the uncertainty inherent in network studies. All code used in the paper is made available in the companion supplementary material.
